---
title: "Course 1: Reproducible Research"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=15, fig.height=8,
                      echo=TRUE,eval=TRUE,options(scipen = 1, digits = 2))
```

```{r}
activity <- read.delim("activity.csv", header=TRUE, sep=",", stringsAsFactors=FALSE, dec=".")

activity$date<-as.Date(activity$date, "%Y-%m-%d")

activityClean<-activity[!is.na(activity$steps),]
```

## What is mean total number of steps taken per day?

1.Number of steps taken per day.

```{r echo=TRUE,message=FALSE,results='hide',fig.keep='all'}
library(ggplot2)
library(dplyr)
# Sum for each month
stepsPerDay <- activityClean
stepsPerDay <-
stepsPerDay %>% group_by(date) %>% summarise(sum_steps = sum(steps, na.rm =
TRUE))

#Finding the mean and median
mean<-round(mean(stepsPerDay$sum_steps),0)
median<-round(median(stepsPerDay$sum_steps),0)

# Plotting the steps per day

p1 <- ggplot(stepsPerDay, aes(x = sum_steps)) + labs(title = "Number of Steps Per Day") +
labs(x = "Number of Steps", y = "Count")  +   geom_histogram(binwidth = 400, col = "darkblue",fill = "gray") +
geom_vline(aes(xintercept = mean(stepsPerDay$sum_steps),color = "mean"),linetype = "dashed",size = 1) +
geom_vline(aes(xintercept = median(stepsPerDay$sum_steps),color = "median"),linetype = "dotted",size = 1)+
scale_color_manual(name = "statistics",values = c(median = "#00AFBB", mean = "#E7B800"))

p1

```
The mean steps per day are `r mean` and the median steps per day are `r median`.

## What is the average daily activity pattern?

1.Average of number of steps taken per interval.


```{r echo=TRUE,message=FALSE}
library(ggplot2)
library(dplyr)
# Mean for each month
meanStepsPerDay <- activityClean
meanStepsPerDay <-
meanStepsPerDay %>% group_by(interval) %>% summarise(mean_steps = mean(steps, na.rm =
TRUE))

# Max Steps
maxSteps<-max(meanStepsPerDay$mean_steps, na.rm=TRUE)
meanInterval<-meanStepsPerDay[meanStepsPerDay$mean_steps==maxSteps,1]

# Plotting the average steps per day

p2 <- ggplot(meanStepsPerDay, aes(x = interval, y=mean_steps)) +
geom_line(na.rm=TRUE,color="#00AFBB") + labs(title = "Average Number of Steps Per Day") + labs(x = "Interval", y = "Number of Steps") + geom_hline(aes(yintercept = max(meanStepsPerDay$mean_steps, na.rm=TRUE),color = "max"),linetype = "dashed", size = 1) + scale_color_manual(name = "statistics",values = c(max = "#E7B800"))

p2

```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is `r meanInterval`.

## Imputing missing values

1.What are the total number of rows with missing values in the dataset?


```{r echo=TRUE,message=FALSE}
library(ggplot2)
library(dplyr)
# Number of rows with missing values
rowNA<-sum(apply(is.na(activity), 1, sum))

```
The total number of rows with NAs is `r rowNA`.

2.Filling in all of the missing values in the dataset


```{r echo=TRUE,message=FALSE}
library(ggplot2)
library(dplyr)
# Mean for each month
meanStepsPerDay <- activityClean
meanStepsPerDay <-
meanStepsPerDay %>% group_by(interval) %>% summarise(mean_steps = mean(steps, na.rm =
TRUE))

#Assign the number of steps of the mean for that interval.
newAct<-activity
newActAvg<-merge(x=newAct,y=meanStepsPerDay, by="interval", all.x=TRUE)

newActAvg$steps <- ifelse(is.na(newActAvg$steps), round(newActAvg$mean_steps,1), round(newActAvg$steps,1))

#Drop the mean_steps column
newActAvg$mean_steps<-NULL

```
3.Histogram of the total number of steps taken each day.

```{r echo=TRUE,message=FALSE,results='hide',fig.keep='all'}
library(ggplot2)
library(dplyr)
# Sum for each month
newActAvgSum <-
newActAvg %>% group_by(date) %>% summarise(sum_steps = sum(steps, na.rm =
TRUE))

#Finding the mean and median
meanNA<-round(mean(newActAvgSum$sum_steps),0)
medianNA<-round(median(newActAvgSum$sum_steps),0)

# Plotting the steps per day

p1 <- ggplot(newActAvgSum, aes(x = sum_steps)) + labs(title = "Number of Steps Per Day") +
labs(x = "Number of Steps", y = "Count")  +   geom_histogram(binwidth = 400, col = "darkblue",fill = "gray") +
geom_vline(aes(xintercept = mean(stepsPerDay$sum_steps),color = "mean"),linetype = "dashed",size = 1) +
geom_vline(aes(xintercept = median(stepsPerDay$sum_steps),color = "median"),linetype = "dotted",size = 1)+
scale_color_manual(name = "statistics",values = c(median = "#00AFBB", mean = "#E7B800"))

p1

```

The mean steps per day are `r meanNA` and the median steps per day are `r medianNA`. The difference is barely noticeable with the original calculation.

## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE,message=FALSE,results='hide',fig.keep='all'}
library(ggplot2)
library(dplyr)

# Calculating if days are weekdays or not
newActAvg$weekday<-NA
newActAvg$weekday<-ifelse(weekdays(newActAvg$date) == "Saturday" | weekdays(newActAvg$date) == "Sunday" ,"Weekend","Weekday")


# Mean for each month
meanStepsWeekday <- newActAvg
meanStepsWeekday <-
meanStepsWeekday %>% group_by(interval,weekday) %>% summarise(mean_steps = mean(steps, na.rm =
TRUE))


# Plotting the average steps per day

p2 <- ggplot(meanStepsWeekday, aes(x = interval, y=mean_steps)) +
geom_line(color="#00AFBB") + facet_wrap(~weekday, ncol = 1, nrow=2)  + labs(title = "Average Number of Steps being Weekend or Weekday") + labs(x = "Interval", y = "Number of Steps") 

p2

```

The mean steps per day are `r meanNA` and the median steps per day are `r medianNA`. The difference is barely noticeable with the original calculation.

```{r echo=TRUE,message=FALSE,results='hide',fig.keep='all'}
#library("knitr")
#knit("PA1_template.Rmd")
```

